Response to the defense feedback

### 2021-06-03 Chapter-3/Chapter-3.tex
Updates from Overleaf

DATA~\cite{217537} identifies
side-channel leakages by finding differences in execution traces of
the test program under \emph{various secret inputs}. According to
the original DATA paper, it uses 443 different traces to analyze the
side-channel vulnerabilities in symmetric cyphers and 450 different
traces to analyze the side-channel vulnerabilities in asymmetric
cyphers. On the other hand, \detect{} detects side-channel
vulnerabilities from one execution trace. \detect{} uses symbolic
analysis to extract formulas that model each side-channel
leakage. After that, we sample the formula with \emph{various
  secret inputs} to detect and quantify each leakage site. In
theory, DATA might have better code coverage than \detect{} because it
uses more execution traces,  but \detect{} has the following
advantages. a) \detect{} is faster than DATA. For example, it takes
116 minutes for DATA to detect vulnerabilities in the RSA implementation in OpenSSL 1.1.0f\@. \detect{} only spends 34 minutes, as shown in Table~\ref{table:over_result}. It takes 13 minutes and 20 minutes
for DATA to analyze the side-channel leakages in AES and DES,
respectively. On the other hand, \detect{} finishes its analysis in
less than ten seconds while finding all the leakages reported by
DATA. b) Because \detect{} does not exexute the test program again when we
have a new \emph{secret input}, \detect{} can test more input secrets
on these formulas within the same time to achieve better precision.
c) DATA tries to use leakage models (domain knowledge) to classify each leakage.
The strength of \ctool{} is that it does not need such domain knowledge.
DATA reports 278 control-flow and 460 memory-access leaks for the RSA implementation in OpenSSL 1.1.0f. Among these leakages, they find one new vulnerability in RSA after some manual analysis.
\detect{} finds the vulnerability.  For each leakage site, \detect{} can provide
concrete examples to trigger the issue.
\subsection{Case Studies}
\subsubsection*{DES and AES}
Our tool confirms that both implementations in OpenSSL and mbedTLS have side-channel leakages
sites. Moreover, we find that all the leakages belong to the secret-dependent data accesses. The reference implementation of AES uses four lookup tables to speed up the computation. However, such implementation are vulnerable to side-channel attacks. Previous work has shown an end-to-end attack to fully recover the key. Our findings is similar to the previous work. We also notice that later version of OpenSSL (after 1.0.1) uses a modified version with smaller S-table. However, \detect{} also find the leakage sites in the version. Recent work also finds the sites through theoretical analysis.

\subsubsection*{RSA}
We also evaluates our tool on the RSA implementations. With the optimizations
introduced in the previous section, we do not apply any domain knowledge to
simplify the analysis. Therefore, our tools can identify all the leakage sites
reported by CacheD~\cite{203878} in a shorter time. Our tool finds that most
leakages in RSA occur in the big number implementation. We also find the newer
versions of RSA in OpenSSL tend to have fewer leakages detected by \detect{}.

@@

\begin{figure}[ht]
\centering
\begin{lstlisting}[xleftmargin=.02\textwidth,xrightmargin=.01\textwidth]
while (!BN_is_bit_set(B, shift)) { /* note that 0 < B */
    shift++;
    if (BN_is_odd(X)) {
        if (!BN_uadd(X, X, n))
            goto err;
    }
    ...
    if (!BN_rshift1(X, X))  // It causes the leak severe!
        goto err;
    ...
}
\end{lstlisting}
    \caption{Unknown sensitive secret-dependent branch leaks from function 
             \textsf{int\_bn\_mod\_inverse} in OpenSSL 1.1.1. The function can leak the last digit from big number \textbf{X}. However, the leak is more severe because of the 
             function \textsf{BN\_rshift1}. Each time function \textsf{BN\_rshift1}
             will shift \textbf{X} right by one and places the result in \textbf{X}. Therefore,
             an attacker can infer multiple bits of \textbf{X} by observing the branch at line 3.}
    \label{chapter3:fig:unknown_leakage}
\end{figure}

\subsubsection*{Monocypher}\label{eval:mono}
Monocypher is a small, easy to use cryptographic library with
performance comparable to LibSodium~\cite{libsodium} and NaCl~\cite{bernstein2012security}.
We choose four ciphers that are
designed to be side-channel resistant from the library.
Because these ciphers have no
data flow from secrets to branch conditions and load addresses.
Monocypher should be safe under our threat models.
We analyze these ciphers with \detect{}, and it reports no leaks.
This indicates that \detect{} is effective for validating countermeasures.

>>>>

DATA~\cite{217537} identifies
side-channel leakages by finding differences in execution traces of
the test program under \emph{various secret inputs}. According to
the original DATA paper, it uses 443 different traces to analyze the
side-channel vulnerabilities in symmetric cyphers and 450 different
traces to analyze the side-channel vulnerabilities in asymmetric
cyphers. On the other hand, \detect{} detects side-channel
vulnerabilities from one execution trace. \detect{} uses symbolic
analysis to extract formulas that model each side-channel
leakage. After that, we sample the formula with \emph{various
  secret inputs} to detect and quantify each leakage site. In
theory, DATA might have better code coverage than \detect{} because it
uses more execution traces,  but \detect{} has the following
advantages. a) \detect{} is faster than DATA. For example, it takes
116 minutes for DATA to detect vulnerabilities in the RSA implementation in OpenSSL 1.1.0f\@. \detect{} only spends 34 minutes, as shown in Table~\ref{table:over_result}. It takes 13 minutes and 20 minutes
for DATA to analyze the side-channel leakages in AES and DES,
respectively. On the other hand, \detect{} finishes its analysis in
less than ten seconds while finding all the leakages reported by
DATA. b) Because \detect{} does not exexute the test program again when we
have a new \emph{secret input}, \detect{} can test more input secrets
on these formulas within the same time to achieve better precision.
c) DATA tries to use leakage models (domain knowledge) to classify each leakage.
The strength of \ctool{} is that it does not need such domain knowledge.
DATA reports 278 control-flow and 460 memory-access leaks for the RSA implementation in OpenSSL 1.1.0f. Among these leakages, they find one new vulnerability in RSA after some manual analysis.
\detect{} finds the vulnerability.  For each leakage site, \detect{} can provide
concrete examples to trigger the issue.
\subsection{Case Studies}
\subsubsection*{DES and AES}
Our tool confirms that both implementations in OpenSSL and mbedTLS have side-channel leakages
sites. Moreover, we find that all the leakages belong to the secret-dependent data accesses. The reference implementation of AES uses four lookup tables to speed up the computation. However, such implementation are vulnerable to side-channel attacks. Previous work has shown an end-to-end attack to fully recover the key. Our findings is similar to the previous work. We also notice that later version of OpenSSL (after 1.0.1) uses a modified version with smaller S-table. However, \detect{} also finds leakage sites in the version. Recent work also finds the sites through theoretical analysis.

\subsubsection*{RSA}
We also evaluate our tool on the RSA implementations. With the optimizations
introduced in the previous section, we do not apply any domain knowledge to
simplify the analysis. Therefore, our tools can identify all the leakage sites
reported by CacheD~\cite{203878} in a shorter time. Our tool finds that most
leakages in RSA occur in the big number implementation. We also find the newer
versions of RSA in OpenSSL tend to have fewer leakages.

@@

\begin{figure}[ht]
\centering
\begin{lstlisting}[xleftmargin=.02\textwidth,xrightmargin=.01\textwidth]
while (!BN_is_bit_set(B, shift)) { /* note that 0 < B */
    shift++;
    if (BN_is_odd(X)) {
        if (!BN_uadd(X, X, n))
            goto err;
    }
    ...
    if (!BN_rshift1(X, X))  // It causes the leak severe!
        goto err;
    ...
}
\end{lstlisting}
    \caption{Previously unknown sensitive secret-dependent branch leaks from function 
             \textsf{int\_bn\_mod\_inverse} in OpenSSL 1.1.1. The function can leak the last digit from big number \textbf{X}. However, the leak is more severe because of the 
             function \textsf{BN\_rshift1}. Each time, function \textsf{BN\_rshift1}
             shifts \textbf{X} right by one and places the result in \textbf{X}. Therefore,
             an attacker can infer multiple bits of \textbf{X} by observing the branch at line 3.}
    \label{chapter3:fig:unknown_leakage}
\end{figure}

\subsubsection*{Monocypher}\label{eval:mono}
Monocypher is a small, easy to use cryptographic library with
performance comparable to LibSodium~\cite{libsodium} and NaCl~\cite{bernstein2012security}.
We choose four ciphers that are
designed to be side-channel resistant from the library.
These ciphers have no
data flow from secrets to branch conditions and load addresses.
Monocypher should be safe under our threat models.
We analyze these ciphers with \detect{}, and it reports no leaks.
This indicates that \detect{} is effective for validating countermeasures.
